### 2019-05-23 docs/thesis-qualification/chapters/4_tensor_dl/tensor_dl.tex
prepare for thesis writing

Dictionary learning is related to CS but seeks to obtain compact and meaningful signal representations, known as dictionary, from learning algorithms or models. DL is a signal processing technique for sparse representation of signals as basis vectors, learning the representations from training data, as dictionaries. DL is based on the principle that some observations can be described by a sparse subset of atoms taken from a redundant dictionary, which represents the causes of some observations of the world. The sparse representation in terms of such dictionaries has attracted increased interest for compressive sensing and for solving problems such as denoising, compression, image processing, data decomposition, feature extraction and classification, when the learning includes a class separability criteria in the objective function \cite{tosic2011dictionary, zhang2010discriminative, zhu2016coupled,ravishankar2011mr}.

There are two major approaches for DL. First is the analytic approach, in which Discrete Cosine Transform basis, wavelets, curvelets and other nonadaptive functions are used as atoms to construct the dictionaries. Second is the learning-based approaches, such as the unsupervised learning for dictionary construction [9] and the online dictionary learning [11], [10], which use machine learning methods to construct the dictionary.

In the analytic approach, some pre-defined functions are used to construct the dictionary. Curvelets [24], which tracks the shape of the discontinuity set, offers efficient and near-optimal representation of smooth objects. Shearlets [25], which is obtained from dilations, action of translations and shear transformations, has nice geometric properties and mathematical properties for image representation. Bandelets [26], which specifies the geometry as a vector field, is designed to improve the image compression and noise reduction performance.

In the learning-based approach, machine learning methods are used to construct the dictionary from the training data. The least square error is used by the method of the optimal directions (MOD) [27] to update the dictionary iteratively. While analytic dictionaries permit to capture the global structure of a signal and allow a fast implementation, learned dictionaries often perform better in applications as they are more adapted to the considered class of signals.

@@

Multidimensional parameter estimation and learning multidimensional separable dictionaries are growing research problems. The crucial idea about separable dictionaries is to allow the dictionary to have a separable structure, where separable means that the dictionary $\textbf{D}$ is given by the Kronecker product of two smaller dictionaries $\boldsymbol{A}^{(1)} \in \mathbb{R}^{h \times a}$ and $\boldsymbol{A}^{(2)} \in \mathbb{R}^{w \times b}$, for example. Roemer \emph{et al.} \cite{roemer2014tensor} show that the multidimensional dictionary estimation problem can be efficiently formulated in terms of tensors, and that their results outperform existing schemes by exploiting the multilinear structure of the problem.

Considering that big data problems require techniques to deal with multidimensional data in order to make sense of structure and relationship of many dimensions, and also considering that a key challenge to use sparse coding and dictionary learning for classification is how to find proper dictionaries and coefficients that highlight discriminative structure and relationships of one dataset, this work propose a tensor-based DL for fraud detection from imbalanced data, in order to apply the tensor decomposition for DL methods to highlight the discriminative sensing of a fraud detection dataset.

This chapter is organized as follows. Section \ref{sec:4_motivation} presents one experiment and results that motivate the proposed work. In Section \ref{sec:4_relatedworks}, related works are discussed. Section \ref{sec:4_datamodel} presents the data model and the evaluated datasets. Section \ref{sec:4_proposal} describes the proposed approach for fraud detection from mobile money transactions. Section \ref{sec:2_experimentalresults} discusses the experimental validation and presents the results, and Section \ref{sec:4_conclusion} draws the conclusions and the suggestions for future work.

@@

Considering that DL aims to obtain meaningful dictionaries, it is possible to evaluate DL algorithms by their performance to generate dictionaries that better represent one target data. Therefore, we can compare DL algorithms regarding recovering of a known dictionary. We conduct a evaluation of tensor-based DL algorithms for recovering a known dictionary in order to analyze the gains of tensor-based approaches in comparison to well known algorithms.

@@

\begin{equation}\label{eq:4_eq01}
	\boldsymbol{X} = \boldsymbol{A} \cdot \boldsymbol{S} + \boldsymbol{W},
\end{equation}

where $\boldsymbol{X} \in \mathbb{C}^{M \times T}$ represents $T$ consecutive observations from $M$ features, $\boldsymbol{A} \in \mathbb{C}^{M \times N}$ is the overcomplete dictionary where $N \ll T$, $\boldsymbol{S} \in \mathbb{C}^{N \times T}$ represents the sparse coefficient matrix, $\boldsymbol{W} \in \mathbb{C}^{M \times T}$ is the additive noise, and $M < N < T$.

@@

\begin{equation}\label{eq:4_eq02}
	\boldsymbol{X} = (\boldsymbol{A}^{(1)} \otimes \boldsymbol{A}^{(2)}) \cdot \boldsymbol{S} + \boldsymbol{W},
\end{equation}

where $\boldsymbol{A} = (\boldsymbol{A}^{(1)} \otimes \boldsymbol{A}^{(2)})$, $\boldsymbol{A}^{(1)} \in \mathbb{C}^{M_1 \times N_1}$, $\boldsymbol{A}^{(2)} \in \mathbb{C}^{M_2 \times N_2}$, $M = M_1 \times M_2$, and $N = N_1 \times N_2$.

In order to evaluate tensor-based DL algorithms for recovering a known dictionary, we generate two random dictionaries $\boldsymbol{A}^{(1)}$ and $\boldsymbol{A}^{(2)}$ from an i.i.d. zero mean Gaussian random process, and calculate the dictionary $\boldsymbol{A}$ through the Kronecker product, according to

\begin{equation}\label{eq:4_eq03}
	\boldsymbol{A} = (\boldsymbol{A}^{(1)} \otimes \boldsymbol{A}^{(2)}).
\end{equation}

Therefore, $\boldsymbol{A}$ is the known dictionary that shall be used for recovery evaluation.

Subsequently, we generate a synthetic data set $\boldsymbol{X}$ using the given dictionary $\boldsymbol{A}$, according to Equation \ref{eq:4_eq01}, where $\boldsymbol{S}$ is a random sparse coefficient matrix for what we assume that each column has $K = 5$ non-zero entries, and $\boldsymbol{W}$ is the additive white gaussian noise. Thus, we estimate the initial $\boldsymbol{\hat{A}}$ dictionary from a normalized subset of $\boldsymbol{X}$ and make the dictionary separable decomposing the approximation of $\boldsymbol{\hat{A}}$ and generating $\boldsymbol{\hat{A}}^{(1)}$ and $\boldsymbol{\hat{A}}^{(2)}$ according to \cite{van1993approximation}. Afterward, we update $\boldsymbol{\hat{A}}$ through

\begin{equation}\label{eq:4_eq04}
	\boldsymbol{\hat{A}} = (\boldsymbol{\hat{A}}^{(1)} \otimes \boldsymbol{\hat{A}}^{(2)}),
\end{equation}

and use it for DL experiments using MOD \cite{engan1999method}, K-SVD \cite{aharon2006rm}, RLS-DLA \cite{skretting2010recursive}, T-MOD \cite{roemer2014tensor} and K-HOSVD \cite{roemer2014tensor}. Due to its low computational complexity, we employ the Orthogonal Matching Pursuit (OMP) algorithm\cite{davis1997adaptive} for the sparse recovery stage in all the dictionary learning algorithms, assuming that K is known. However, other solver for sparse recovery can be used, such as Basis Pursuit \cite{chen2001atomic}, the LASSO algorithm \cite{tibshirani1996regression} or greedy \cite{davis1997adaptive}.

The final estimated dictionaries $\boldsymbol{\hat{A}}$ recovered by each DL algorithm are compared to the known dictionary $\boldsymbol{A}$, through a measure of the distance between two dictionaries. The recovered dictionary is then compared to the known dictionary, by using the angle between a known atom and a recovered atom as a distance measure. The number of identified atoms is computed by comparing these angles to a limit of $\beta_{lim} = 8.11$ degrees, that corresponds to the case where the inner product of the two compared atoms is equal to 0.99, and where each atom has 2-norm equal to 1.

For this experiments we adopt an additive noise with $snr = 20$ and 100 iterations for dictionary learning. All experiments run 100 trials in order to obtain confident results.

This evaluation extends Roemer \emph{et al.} \cite{roemer2014tensor} including RLS-DLA for comparison, and for adding the evaluation of the degree distribution for atom identification and the evaluation of the atom identification according to the dictionary size variation. Additionally, we conduct a experimental evaluation, while Roemer \emph{et al.} \cite{roemer2014tensor} presents numerical results obtained via Monte Carlo simulations.

@@

Figure \ref{fig:fig1} presents the results for recovering a dictionary of $M=20$ and $N=50$. We see that the RLS-DLA method performs better than other methods, although T-MOD presents the best results for identification of up to 15 atoms of 50. It is possible to observe that T-MOD and K-HOSVD are better than MOD and K-SVD, and that the tensor-based methods outperforms MOD and K-SVD significantly for this experiment. Finally, Figure \ref{fig:fig1} shows that RLS-DLA recovered almost 40 atoms in $\beta_{lim} = 8.11$ degrees, while other methods present results lower results.

\begin{figure}[!htb]
	\centering 
	\includegraphics[width=10cm]{figures/5_20_2000_1000_100.eps}
	\caption{Cumulative atom identification per degree rates. $T=2000$, $M=20$, $N=50$}
	\label{fig:fig1}
\end{figure}

Figure \ref{fig:fig2} presents the results for recovering a dictionary of $M=80$ and $N=30$. It is possible to observe that the tensor-base methods outperform MOD, RLS-DLA and K-SVD in number of identified atoms and the required degrees for identification. Note that MOD performs quite similar to RLS-DLA, while K-SVD is the worst evaluated case for this experiment. Considering the tensor-based methods, T-MOD outperforms K-HOSVD with higher atom identification over lower degrees. The experimental results verify that the tensor-based dictionary learning algorithms are able to identify more dictionary atoms when dealing with larger dictionaries and that they converge faster.

\begin{figure}[!htb]
	\centering 
	\includegraphics[width=10cm]{figures/5_20_2000_24000_100.eps}
	\caption{Cumulative atom identification per degree rates. $T=2000$, $M=80$, $N=300$}
	\label{fig:fig2}
\end{figure}

The results reveals that tensor-based algorithms can perform better for dictionary recovery problems when the dictionary size increases. Since dictionary learning and reconstruction error are usually used for sparse representation classification (SRC) problems, it is possible to suppose that tensor-based algorithms can improve results of classification problems, as well as can extract hidden patterns and structure in multidimensional data analytics problems.

Recent developments in science and technology have enabled the growth and availability of raw data to occur at an explosive rate. This has created an immense opportunity for knowledge discovery and data engineering research to play an essential role in a wide range of applications from daily civilian life to national security. However, the high availability of raw data increases the challenges related to big data analytics and to imbalanced data, which corresponds to data sets exhibiting significant imbalances of classes or rare events of some classes. The fundamental issue with the imbalanced learning problem is the ability of imbalanced data to significantly compromise the performance of most standard learning algorithms

A key challenge to use SC and DL for classification is how to find proper dictionaries and coefficients that highlight discriminative structure and relationships of one dataset. Therefore, we introduce a tensor-based DL method for fraud detection from unbalanced data. Specifically, we apply a sparse representation based classification (SRC) method through learning a tensor-based dictionary and evaluate the reconstruction error for fraud classification.

@@

One of the key principles of dictionary learning is that the dictionary has to be inferred from the input data. The emergence of sparse dictionary learning methods was stimulated by the fact that in signal processing one typically wants to represent the input data using as few components as possible. Before this approach the general practice was to use predefined dictionaries (such as Fourier or wavelet transforms). However, in certain cases a dictionary that is trained to fit the input data can significantly improve the sparsity, which has applications in data decomposition, compression and analysis and has been used in the fields of image denoising and classification, video and audio processing. Sparsity and overcomplete dictionaries have immense applications in image compression, image fusion and inpainting.

MOD \cite{engan1999method} is a DL algorithm that is commonly used for extensions of Least-Squares (LS) type schemes, also known as iterative least squares dictionary learning algorithm (ILS-DLA) \cite{engan2007family}. MOD defines that if we know $\boldsymbol{S}$ in equation \ref{eq:4_eq01}, $\boldsymbol{A}$ can be estimated solving a linear LS problem. Additionally, knowing $\boldsymbol{A}$ we obtain $\boldsymbol{S}$ by solving a sparse recovery problem. MOD iterates between the two steps, starting with an initial guess of $\boldsymbol{A}$, denoted as $\hat{\boldsymbol{A}}$, which can be estimated, generated by previous dictionary learning or generated analytically. Despite - simplicity, the MOD algorithm has been criticized for its slow convergence, which is due to the fact that $\boldsymbol{A}$ and $\boldsymbol{S}$ are updated separately, which may lead to a slow convergence \cite{aharon2006rm}.

K-SVD \cite{aharon2006rm} is a generalization of the K-means algorithm for dictionary learning. After the sparse approximation step, the dictionary update is performed by sequentially updating each column of $\boldsymbol{A}$ using SVD to minimize the approximation error. The update step is hence a generalized K-means algorithm since each patch can be represented by multiple atoms and with different weights. In practice, dictionaries learned with K-SVD
have shown excellent performance in image denoising.

@@

Dimensionality reduction can first be achieved by selecting a subset of functions from a large, fixed dictionary that is used for the analysis of particular signals. These functions then determine a subspace of reduced dimension, where classification can be performed by computing the nearest neighbor points among the projected data. A simple method to build such a subspace consists in modifying the sparse approximation methods such that the objective function is augmented with a discrimination term that represents the separability properties of the projection subspace. This kind of approach typically tries to maximize the variance between the active atoms from $\boldsymbol{A}$ that represent signals in different classes. The reduced subspace used for classification is finally formed by the subset of atoms in $\boldsymbol{A}$ whose corresponding coefficients in $\boldsymbol{S}$ are nonzero. The subset selection problem can be interpreted as the inference step in the dictionary learning methods when the objective function is modified to include a discriminative term \cite{tosic2011dictionary}.

An important advantage of redundant dictionaries for classification is that signal analysis can be performed with functions that are likely to match the data characteristics in different classes of signals. Similarly to data approximation problems, data analysis applications can further benefit from dictionary learning methods. The previous section describes subspace selection methods from predefined dictionaries. DL algorithms can improve the classification performance, as it leads to a better adaptation of the dictionary by enforcing sparsity in the representation of data in the different classes. The atoms in a dictionary $\boldsymbol{A}$ that is computed with dictionary learning methods generally capture the most important constitutive components of the signals. They naturally permit to classify the data into the corresponding linear subspace. However, there is no guarantee that the subspace built on a learned dictionary $\boldsymbol{A}$ is truly optimal for classification, as it targets efficient representation but not necessarily class separability \cite{tosic2011dictionary}.

@@

Considering that tensor-based algorithms can perform better for dictionary recovery problems when the dictionary size increases, according to subsection \ref{sec:4_motivation_results}, we believe that tensor-based DL algorithms can improve the discriminative sensing for classification classification problems. Therefore, we propose use tensor-based DL for learning fraudulent and legitimate data separately, and apply the learned dictionaries to reconstruct a test signal and classify it as fraud or legitimate, according to the minimum reconstruction error measured by some metrics

@@

We adopt a synthetic financial datasets for fraud detection\footnote{https://www.kaggle.com/ntnu-testimon/paysim1} in order to validate our proposal. This dataset is synthetically generated by  PaySim \cite{lopez2016paysim}, which is a financial simulator that simulates mobile money transactions based on an original dataset and provides a synthetic dataset as an approach to such a problem. PaySim uses aggregated data from the private dataset to generate a synthetic dataset that resembles the normal operation of transactions and injects malicious behavior to later evaluate the performance of fraud detection methods. PaySim simulates mobile money transactions based on a sample of real transactions extracted from one month of financial logs from a mobile money service implemented in an African country. The original logs were provided by a multinational company

The selected dataset is composed by 9 features (variables) and 2.770.408 transactions (observations). It is also provided one label that classifies each transaction as fraudulent or not. As the majority dataset of anomaly or fraud detection, the dataset of mobile money transactions is imbalanced, where the number of fraudulent transactions is 8.213 against 2.762.196 of normal transactions. Therefore, the fraud class represents 0,30 \% of the total transactions.

For evaluation of the proposed classification approach we adopt a data splitting into training and test data, where the training data has 70 \% of the whole dataset. We denote the training data as $X_r$ and the test data as $X_e$. The dataset can also be divided by normal and fraudulent transactions, where the normal train data is denoted as $X_r^o$ and the fraud train data is denoted as $X_r^f$. The same notation of fraud, normal, train and test is applied to qualify $A$ and $S$.

@@

% \begin{equation}\label{eq:4_eq04}
% \boldsymbol{\mathcal{X}} = \boldsymbol{\mathcal{S}} \times_1 \boldsymbol{A}^{(1)} \times_2 \boldsymbol{A}^{(2)} + \boldsymbol{\mathcal{W}},
% \end{equation}

% where $\boldsymbol{\mathcal{X}} \in \mathbb{C}^{M_1 \times M_2 \times T}$, $\boldsymbol{\mathcal{S}} \in \mathbb{C}^{N_1 \times N_2 \times T}$, and $\boldsymbol{\mathcal{W}} \in \mathbb{C}^{M_1 \times M_2 \times T}$ are rearranged versions of the matrices $\boldsymbol{X}$, $\boldsymbol{S}$, and $\boldsymbol{W}$ such that $\boldsymbol{X} = [\boldsymbol{\mathcal{X}}]_{(3)}^T$, $\boldsymbol{S} = [\boldsymbol{\mathcal{S}}]_{(3)}^T$, and $\boldsymbol{W} = [\boldsymbol{\mathcal{W}}]_{(3)}^T$, respectively.

% We consider a randomly generated data in order to evaluate the proposed approach, where initially $\boldsymbol{A}^{(1)}$ and $\boldsymbol{A}^{(2)}$ are gaussian distributed and randomly generated, then the dictionary $\boldsymbol{A}$ is obtained. In order to be able to evaluate the dictionary reconstruction, the sparse data $\boldsymbol{\mathcal{X}}$ is generated from the dictionary $\boldsymbol{A}$, which is used for further comparisons. We adopt 20 as the signal-to-noise ratio (snr) and 5 as the sparseness factor of the data $\boldsymbol{\mathcal{X}}$.

% The dictionary estimation is evaluated by applying the proposed approach to estimate the dictionary $\boldsymbol{A}$ from the sparse data $\boldsymbol{\mathcal{X}}$, according the selected number of sample training and iterations for the optimization problem.

@@

We propose a Tensor-based Sparse Representation Classification approach to use tensor-based DL for learning fraudulent and legitimate dictionaries separately, and apply the learned dictionaries to reconstruct a test signal and classify it as fraud or legitimate, according to the minimum reconstruction error measured by some metrics.

>>>>

Dictionary Learning (DL) is related to CS but seeks to obtain compact and meaningful signal representations, known as dictionary, from learning algorithms or analytical models. DL is a signal processing technique for sparse representation of signals as basis vectors, learning the representations from training data, as dictionaries. DL is based on the principle that some observations can be described by a sparse subset of atoms taken from a redundant dictionary, which represents the causes of some observations of the world. The sparse representation in terms of such dictionaries has attracted increased interest for compressive sensing and for solving problems such as denoising, compression, image processing, data decomposition, feature extraction and classification, when the learning includes a class separability criteria in the objective function \cite{tosic2011dictionary, zhang2010discriminative, zhu2016coupled,ravishankar2011mr}.

There are two major approaches for DL. First is the analytic approach, in which Discrete Cosine Transform basis, wavelets, curvelets and other nonadaptive functions are used as atoms to construct the dictionaries. Second is the learning-based approaches, such as the unsupervised learning for dictionary construction [9] and the online dictionary learning \cite{mairal2009online}, which use machine learning methods to construct the dictionary. Learning-based approaches usually uses least square error by the method of the optimal directions (MOD) \cite{engan1999method} to update the dictionary iteratively. While analytic dictionaries permit to capture the global structure of a signal and allow a fast implementation, learned dictionaries often perform better in applications as they are more adapted to the considered class of signals.

@@

Multidimensional parameter estimation and learning multidimensional separable dictionaries are growing research problems. The crucial idea about separable dictionaries is to allow the dictionary to have a separable structure, where separable means that the dictionary $\textbf{A}$ is given by the Kronecker product of two smaller dictionaries $\textbf{A}^{(1)} \in \mathbb{R}^{h \times a}$ and $\textbf{A}^{(2)} \in \mathbb{R}^{w \times b}$, for example. Roemer \emph{et al.} \cite{roemer2014tensor} show that the multidimensional dictionary estimation problem can be efficiently formulated in terms of tensors, and that their results outperform existing schemes by exploiting the multilinear structure of the problem.

Considering that big data problems require techniques to deal with multidimensional data in order to make sense of structure and relationship of many dimensions, and also considering that a key challenge to use sparse coding and dictionary learning for classification is how to find proper dictionaries and coefficients that highlight the discriminative structure and relationships of one dataset, this work proposes a tensor-based DL for fraud detection from imbalanced data, in order to apply the tensor decomposition for DL methods to highlight the discriminative sensing for fraud detection from a dataset of mobile money transactions.

This chapter is organized as follows. Section \ref{sec:4_motivation} presents one experiment and results that motivate the proposed work. In Section \ref{sec:4_relatedworks}, related works are discussed. Section \ref{sec:4_datamodel} presents the data model and the evaluated datasets. Section \ref{sec:4_proposal} describes the proposed approach for fraud detection from mobile money transactions. Section \ref{sec:4_experiments} discusses the experimental validation, Section \ref{sec:4_results} presents the results, and Section \ref{sec:4_conclusion} draws the conclusions and the suggestions for future work.

@@

Considering that DL aims to obtain meaningful dictionaries, it is possible to evaluate DL algorithms by their performance to generate dictionaries that better represent one target data. Therefore, we can compare DL algorithms and tensor-based DL algorithms regarding recovering of a known dictionary.

In this section we conduct a evaluation of tensor-based DL algorithms for recovering a known separable dictionary in order to analyze the gains of tensor-based approaches in comparison to well known algorithms.

@@

\begin{equation}\label{eq:4_eq01}
	\textbf{X} = \textbf{A} \cdot \textbf{S} + \textbf{W},
\end{equation}

where $\textbf{X} \in \mathbb{C}^{M \times T}$ represents $T$ consecutive observations from $M$ features, $\textbf{A} \in \mathbb{C}^{M \times N}$ is the overcomplete dictionary where $N \ll T$, $\textbf{S} \in \mathbb{C}^{N \times T}$ represents the sparse coefficient matrix, $\textbf{W} \in \mathbb{C}^{M \times T}$ is the additive noise, and $M < N < T$.

@@

\begin{equation}\label{eq:4_eq02}
	\textbf{X} = (\textbf{A}^{(1)} \otimes \textbf{A}^{(2)}) \cdot \textbf{S} + \textbf{W},
\end{equation}

where $\textbf{A} = (\textbf{A}^{(1)} \otimes \textbf{A}^{(2)})$, $\textbf{A}^{(1)} \in \mathbb{C}^{M_1 \times N_1}$, $\textbf{A}^{(2)} \in \mathbb{C}^{M_2 \times N_2}$, $M = M_1 \times M_2$, and $N = N_1 \times N_2$.

In order to evaluate tensor-based DL algorithms for recovering a known dictionary, we generate two random dictionaries $\textbf{A}^{(1)}$ and $\textbf{A}^{(2)}$ from an i.i.d. zero mean Gaussian random process, and calculate the dictionary $\textbf{A}$ through the Kronecker product, according to Equation

\begin{equation}\label{eq:4_eq03}
	\textbf{A} = (\textbf{A}^{(1)} \otimes \textbf{A}^{(2)}).
\end{equation}

Therefore, $\textbf{A}$ is the known dictionary that shall be used for recovery evaluation.

Subsequently, we generate a synthetic data set $\textbf{X}$ using the given dictionary $\textbf{A}$, according to Equation \ref{eq:4_eq01}, where $\textbf{S}$ is a random sparse coefficient matrix for what we assume that each column has $K = 5$ non-zero entries, and $\textbf{W}$ is the additive white gaussian noise. Thus, we estimate the initial $\hat{\textbf{A}}$ dictionary from a normalized subset of $\textbf{X}$ and make the dictionary separable decomposing the approximation of $\hat{\textbf{A}}$ and generating $\hat{\textbf{A}}^{(1)}$ and $\hat{\textbf{A}}^{(2)}$ according to \cite{van1993approximation}. Afterward, we update $\hat{\textbf{A}}$ through Equation

\begin{equation}\label{eq:4_eq04}
	\hat{\textbf{A}} = (\hat{\textbf{A}}^{(1)} \otimes \hat{\textbf{A}}^{(2)}),
\end{equation}

and use it for DL recovery experiments using MOD \cite{engan1999method}, K-SVD \cite{aharon2006rm}, RLS-DLA \cite{skretting2010recursive}, T-MOD \cite{roemer2014tensor} and K-HOSVD \cite{roemer2014tensor}.

Due to its low computational complexity, we employ the Orthogonal Matching Pursuit (OMP) algorithm\cite{davis1997adaptive} for the sparse recovery stage in all the dictionary learning algorithms, assuming that $K$ is known. However, other solver for sparse recovery can be used, such as Basis Pursuit \cite{chen2001atomic}, the LASSO algorithm \cite{tibshirani1996regression} or greedy \cite{davis1997adaptive}.

The final estimated dictionaries $\hat{\textbf{A}}$ recovered by each DL algorithm are compared to the known dictionary $\textbf{A}$, through a measure of the distance between two dictionaries. The recovered dictionary is then compared to the known dictionary, by using the angle between a known atom and a recovered atom as a distance measure. The number of identified atoms can be computed by comparing these angles to a limit of $\beta_{lim} = 8.11$ degrees, that corresponds to the case where the inner product of the two compared atoms is equal to 0.99, and where each atom has 2-norm equal to 1.

For this experiments we adopt an additive noise with $snr = 20$ and 100 iterations for dictionary learning. All experiments repeats 100 times in order to obtain confident results.

This evaluation extends Roemer \emph{et al.} \cite{roemer2014tensor} including the comparison to RLS-DLA algorithm, and for adding the evaluation of the degree distribution for atom identification and the evaluation of the atom identification according to the dictionary size variation. Additionally, we conduct a experimental evaluation, while Roemer \emph{et al.} \cite{roemer2014tensor} presents numerical results obtained via Monte Carlo simulations.

@@

Figure \ref{fig:fig1} presents the results for recovering a dictionary of $M=20$ and $N=50$. We see that the RLS-DLA method performs better than other methods, although T-MOD presents the best results for identification of up to 15 atoms of 50. It is possible to observe that T-MOD and K-HOSVD are better than MOD and K-SVD, and that the tensor-based methods outperforms MOD and K-SVD significantly for this experiment. Finally, Figure \ref{fig:fig1} shows that RLS-DLA recovered almost 40 atoms in $\beta_{lim} = 8.11$ degrees, while other methods present lower results.

\begin{figure}[!htb]
	\centering 
	\includegraphics[width=10cm]{figures/ch4/s=5_snr=20_L=2000_noIt=100_N=20_K=50.eps}
	\caption{Cumulative atom identification per degree rates. $T=2000$, $M=20$, $N=50$}
	\label{fig:fig1}
\end{figure}

Figure \ref{fig:fig2} presents the results for recovering a dictionary of $M=80$ and $N=30$. It is possible to observe that the tensor-base methods outperform MOD, RLS-DLA and K-SVD in number of identified atoms and the required degrees for identification.

Note that MOD performs quite similar to RLS-DLA, while K-SVD is the worst evaluated case for this experiment. Considering the tensor-based methods, T-MOD outperforms K-HOSVD with higher atom identification over lower degree rates. The experimental results verify that the tensor-based dictionary learning algorithms are able to identify more dictionary atoms when dealing with larger dictionaries and that they converge faster.

\begin{figure}[!htb]
	\centering 
	\includegraphics[width=10cm]{figures/ch4/5_20_2000_24000_100.eps}
	\caption{Cumulative atom identification per degree rates. $T=2000$, $M=80$, $N=300$}
	\label{fig:fig2}
\end{figure}

The results also reveals that tensor-based algorithms can perform better for dictionary recovery problems when the dictionary size increases. Since dictionary learning and reconstruction error are usually used for sparse representation classification (SRC) problems, it is possible to suppose that tensor-based algorithms can improve results of classification problems, as well as can extract hidden patterns and structure in multidimensional data analytics problems.

Recent developments in science and technology have enabled the growth and availability of raw data to occur at an explosive rate. This has created an immense opportunity for knowledge discovery and data engineering research to play an essential role in a wide range of applications from daily civilian life to national security. However, the high availability of raw data increases the challenges related to big data analytics and to imbalanced data, which corresponds to data sets exhibiting significant imbalances of classes or rare events of some classes. The fundamental issue with the imbalanced learning problem is the ability of imbalanced data to significantly compromise the performance of most standard learning algorithms.

A key challenge to use CS and DL for classification is how to find proper dictionaries and coefficients that highlight the discriminative structure and relationships of one dataset. Therefore, we introduce a tensor-based DL method for fraud detection from imbalanced data. Specifically, we apply a sparse representation based classification (SRC) method through learning a tensor-based dictionary and evaluate the reconstruction error of mobile money transactions for fraud classification.

@@

One of the key principles of dictionary learning is that the dictionary has to be inferred from the input data. The emergence of sparse dictionary learning methods was stimulated by the fact that in signal processing one typically wants to represent the input data using as few components as possible. Before this approach the general practice was to use predefined dictionaries (such as Fourier or wavelet transforms). However, in certain cases a dictionary that is trained to fit the input data can significantly improve the sparsity, which has applications in data decomposition, compression and has been used in the fields of image denoising, video and audio processing.

MOD \cite{engan1999method} is a DL algorithm that is commonly used for extensions of Least-Squares (LS) type schemes, also known as iterative least squares dictionary learning algorithm (ILS-DLA) \cite{engan2007family}. MOD defines that if we know $\textbf{S}$ in Equation \ref{eq:4_eq01}, $\textbf{A}$ can be estimated solving a linear LS problem. Additionally, knowing $\textbf{A}$ we obtain $\textbf{S}$ by solving a sparse recovery problem. MOD iterates between the two steps, starting with an initial guess of $\textbf{A}$, denoted as $\hat{\textbf{A}}$, which can be estimated, generated by previous dictionary learning or generated analytically. Despite its simplicity, the MOD algorithm has been criticized for its slow convergence, which is due to the fact that $\textbf{A}$ and $\textbf{S}$ are updated separately, which may lead to a slow convergence \cite{aharon2006rm}.

K-SVD \cite{aharon2006rm} is a generalization of the K-means algorithm for dictionary learning. After the sparse approximation step, the dictionary update is performed by sequentially updating each column of $\textbf{A}$ using SVD to minimize the approximation error. The update step is hence a generalized K-means algorithm since each patch can be represented by multiple atoms and with different weights. In practice, dictionaries learned with K-SVD
have shown excellent performance in image denoising.

@@

Dimensionality reduction can first be achieved by selecting a subset of functions from a large, fixed dictionary that is used for the analysis of particular signals. These functions then determine a subspace of reduced dimension, where classification can be performed by computing the nearest neighbor points among the projected data. A simple method to build such a subspace consists in modifying the sparse approximation methods such that the objective function is augmented with a discrimination term that represents the separability properties of the projection subspace. This kind of classification approach typically tries to maximize the variance between the active atoms from $\textbf{A}$ that represent signals in different classes. The reduced subspace used for classification is finally formed by the subset of atoms in $\textbf{A}$ whose corresponding coefficients in $\textbf{S}$ are nonzero. The subset selection problem can be interpreted as the inference step in the dictionary learning methods when the objective function is modified to include a discriminative term \cite{tosic2011dictionary}.

An important advantage of redundant dictionaries for classification is that signal analysis can be performed with functions that are likely to match the data characteristics in different classes of signals. Similarly to data approximation problems, data analysis applications can further benefit from dictionary learning methods. DL algorithms can improve the classification performance, as it leads to a better adaptation of the dictionary by enforcing sparsity in the representation of data in the different classes. The atoms in a dictionary $\textbf{A}$ that is computed with dictionary learning methods generally capture the most important constitutive components of the signals. They naturally permit to classify the data into the corresponding linear subspace. However, there is no guarantee that the subspace built on a learned dictionary $\textbf{A}$ is truly optimal for classification, as it targets efficient representation but not necessarily class separability \cite{tosic2011dictionary}.

@@

Considering that tensor-based algorithms can perform better for dictionary recovery problems when the dictionary size increases, according to subsection \ref{sec:4_motivation_results}, we believe that tensor-based DL algorithms can improve the discriminative sensing for classification problems. Therefore, we propose use tensor-based DL for learning fraudulent and legitimate data separately, and apply the learned dictionaries to reconstruct a test signal and classify it as fraud or legitimate, according to the minimum reconstruction error measured by cosine similarity.

@@

We adopt a synthetic financial datasets for fraud detection\footnote{https://www.kaggle.com/ntnu-testimon/paysim1} in order to validate our proposal. This dataset is synthetically generated by  PaySim \cite{lopez2016paysim}, which is a financial simulator that simulates mobile money transactions based on an original dataset and provides a synthetic dataset as an approach to such a problem.

PaySim uses aggregated data from the private dataset to generate a synthetic dataset that resembles the normal operation of transactions and injects malicious behavior to later evaluate the performance of fraud detection methods. PaySim simulates mobile money transactions based on a sample of real transactions extracted from one month of financial logs from a mobile money service implemented in an African country. The original logs were provided by a multinational company.

The selected dataset is composed by 9 features (variables) and 2.770.408 transactions (observations). It is also provided one label that classifies each transaction as fraudulent or legitimate. As the majority dataset regarding anomaly or fraud detection, the dataset of mobile money transactions is imbalanced, where the number of fraudulent transactions is 8.213 against 2.762.195 of normal transactions. Therefore, the fraud class represents 0,30 \% of the total transactions.

For evaluation of the proposed classification approach we adopt a data splitting into training and test data, where the training data has 70 \% of the whole dataset. We denote the training data as $\textbf{X}_r$ and the test data as $\textbf{X}_e$. The dataset can also be divided by normal and fraudulent transactions, where the normal train data is denoted as $\textbf{X}_r^o$ and the fraud train data is denoted as $\textbf{X}_r^f$. The same notation of fraud, normal, train and test is applied to qualify $\textbf{A}$ and $\textbf{S}$.

@@

% \begin{equation}\label{eq:4_eq04}
% \textbf{\mathcal{X}} = \textbf{\mathcal{S}} \times_1 \textbf{A}^{(1)} \times_2 \textbf{A}^{(2)} + \textbf{\mathcal{W}},
% \end{equation}

% where $\textbf{\mathcal{X}} \in \mathbb{C}^{M_1 \times M_2 \times T}$, $\textbf{\mathcal{S}} \in \mathbb{C}^{N_1 \times N_2 \times T}$, and $\textbf{\mathcal{W}} \in \mathbb{C}^{M_1 \times M_2 \times T}$ are rearranged versions of the matrices $\textbf{X}$, $\textbf{S}$, and $\textbf{W}$ such that $\textbf{X} = [\textbf{\mathcal{X}}]_{(3)}^T$, $\textbf{S} = [\textbf{\mathcal{S}}]_{(3)}^T$, and $\textbf{W} = [\textbf{\mathcal{W}}]_{(3)}^T$, respectively.

% We consider a randomly generated data in order to evaluate the proposed approach, where initially $\textbf{A}^{(1)}$ and $\textbf{A}^{(2)}$ are gaussian distributed and randomly generated, then the dictionary $\textbf{A}$ is obtained. In order to be able to evaluate the dictionary reconstruction, the sparse data $\textbf{\mathcal{X}}$ is generated from the dictionary $\textbf{A}$, which is used for further comparisons. We adopt 20 as the signal-to-noise ratio (snr) and 5 as the sparseness factor of the data $\textbf{\mathcal{X}}$.

% The dictionary estimation is evaluated by applying the proposed approach to estimate the dictionary $\textbf{A}$ from the sparse data $\textbf{\mathcal{X}}$, according the selected number of sample training and iterations for the optimization problem.

@@

We propose a Tensor-based Sparse Representation Classification approach to use tensor-based DL for learning fraudulent and legitimate dictionaries separately, and apply the learned dictionaries to reconstruct a test signal and classify it as fraud or legitimate, according to the minimum reconstruction error.
